---
title: "ESPN_VAR_Complete_Table"
author: "Oliver Hagger"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R VAR Table Data

```{r message=FALSE, warning=FALSE}
library(rvest)
library(dplyr)
library(stringr)
library(tidyr)
```

## Set Working Directory

```{r echo=FALSE, message=FALSE, warning=FALSE}
setwd("C:/Users/ohagg/OneDrive - University College London/Desktop/VAR_scraping")
```

```{r eval=FALSE, warning=FALSE, include=TRUE}
setwd("C:~~/VAR_scraping")
```

## Define the URL

```{r}

VAR_PAGE_2020_2021 <- "https://www.espn.com/soccer/english-premier-league/story/3929823/how-var-decisions-have-affected-every-premier-league-club"
```

# Read the webpage
```{r}
page <- read_html(VAR_PAGE_2020_2021)
```

## Get team list data

```{r}
team_list <- page %>%
  html_nodes("div.article-body h2") %>%
  html_text() %>%
  gsub("\\s+$", "", .) %>%  # Remove extra spaces at the end
  gsub("[^a-zA-Z]+$", "", .) %>%
  gsub("^Editor's Picks|How to fix VAR.*|The ultimate guide.*|ESPN's .*|Marcotti: .*", "", .) %>%
  unique()

team_list <- team_list[team_list != ""]
```

## Get net score list

```{r}
net_score_list <- page %>%
  html_nodes("div.article-body h2") %>%
  html_text() %>%
  gsub("^.* ", "", .) %>%
  head(20)
```

## Extract general statistics for each team

```{r}

team_stats_list <- page %>%
  html_nodes("div.article-body p") %>%
  html_text() %>%
  grep("Overturns: ", ., value = TRUE)

```

```{r}
# Create a data frame
data_df <- data.frame(
  team_name = team_list,
  net_score = net_score_list,
  stats_combined = team_stats_list
)

```

## Define stats column mapping

```{r}

stats_col_mapping <- list(
  c('overturns_total', 'Overturns'),
  c('overturns_rejected', 'Rejected overturns'),
  c('leading_to_goals_for', 'Leading to goals for'),
  c('leading_to_goals_against', 'Leading to goals against'),
  c('disallowed_goals_for', 'Disallowed goals for'),
  c('disallowed_goals_against', 'Disallowed goals against'),
  c('net_goal_score', 'Net goal score'),
  c('subj_decisions_for', 'Subjective decisions for'),
  c('subj_decisions_against', 'Subjective decisions against'),
  c('net_subjective_score', 'Net subjective score'),
  c('penalties_for', 'Penalties for / against'),
  c('penalties_against', 'Penalties for / against')
)

```

## Create columns

```{r}

stats_col_list <- sapply(stats_col_mapping, `[`, 1)

for (col in stats_col_list) {
  data_df[[col]] <- 0
}

```

## Update columns based on stats combined information

```{r}

for (i in 1:nrow(data_df)) {
  stats_info <- data_df[i, 'stats_combined']
  stats_lines <- strsplit(stats_info, "(?<=\\d)(?=[A-Z])", perl = TRUE)[[1]]
  
  for (line in stats_lines) {
    key <- strsplit(line, ': ')[[1]][1]
    value <- strsplit(line, ': ')[[1]][2]
    
    for (mapping in stats_col_mapping) {
      if (mapping[[2]] == key) {
        data_df[i, mapping[[1]]] <- value
      }
    }
  }
}

```

## Amend penalties_for and penalties_against columns

```{r}

data_df$penalties_for <- str_extract(data_df$penalties_for, "\\d+")
data_df$penalties_against <- str_extract(data_df$penalties_against, "\\d+")

```

## Add year column and drop stats_combined column

```{r}

data_df$year <- '2019/2020'

data_df <- data_df %>%
  select(-stats_combined)

```

```{r}
net_score_columns <- data_df %>% 
  select(starts_with("net_")) %>% 
  names()

data_df <- data_df %>% 
  mutate(across(all_of(net_score_columns), ~ gsub("\\+", "", .)))

```

## Print the resulting dataframe

```{r}

print(data_df)

```

## Export CSV file

```{r}
year <- '2019_2020'
csv_file_name <- paste0("Team", year, ".csv")

write.csv(data_df, csv_file_name, row.names = T)


```

